\documentclass[a4paper,dvipdfmx,10pt,english]{article}
\usepackage{babel}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage[straightquotes,ttzdefault]{newtxtt}
\usepackage{framed}
\usepackage[dvipsnames]{xcolor}
\usepackage{colorframed}
% we put this here but it does not have to be in preamble of
% course
% titled-frame
\colorlet{TFFrameColor}{teal!20}
\colorlet{TFTitleColor}{purple}
\def\colorframedTFconlabcolorcommand{\color{purple}}
% shaded
%\colorlet{shadecolor}{orange!10}
\colorlet{shadecolor}{lightgray!25}

% framed
\FrameRule5pt\relax
\FrameSep9pt\relax % this is framed.sty default anyhow, 3\fboxsep
% attention that \FrameSep influences framed but also shaded, shaded* and
% oframed, which is not so clear from reading only framed documentation
\def\colorframedbordercolorcommand{\color{red!20}}

\title{The \colorframed package%
\vadjust{\vtop to 0pt{\hbox
    to\linewidth{\strut\footnotesize Release 0.9b of
      2022/10/01\hss report issues at \url{https://github.com/jfbu/colorframed/issues}}\vss}}%
}
\author{Jean-François B.}
\date{}
\usepackage{shortvrb}\MakeShortVerb\|
%\usepackage{url}\urlstyle{same}
\usepackage{hyperref}
\def\ctanpackage#1{\href{https://ctan.org/pkg/#1}{#1}}

\usepackage{amssymb}

\usepackage{parskip}
\usepackage{xspace}
\usepackage{metalogo}
\setlogokern{La}{-0.258em}
\setlogokern{aT}{-0.08em}
\setlogokern{eL}{-0.1em}
%\setlogokern{X2}{0em}
\usepackage{enumitem}

\definecolor{joli}{RGB}{225,95,0}
\definecolor{JOLI}{RGB}{225,95,0}
\newcommand\colorframed{%
        \texorpdfstring{{\color{joli}\bfseries colorframed}}{colorframed}\xspace}


% ATTENTION: dvipdfmx syntax here!
{\makeatletter
     \sbox0{\color{RawSienna}%
            \xdef\verbatimpushcolor{\special{color push \current@color}}}%
}
\def\verbatimpopcolor{\special{color pop}}
\AddToHook{env/verbatim/before}{\verbatimpushcolor}
\AddToHook{env/verbatim/end}{\verbatimpopcolor}

\newcommand\ghissue[2][]{\href{https://github.com/jfbu/colorframed/issues/#2}{#1\##2}}

\makeatletter
  %\show\verbatim@font % pas changé at begin doc
  \def\verbatim@font{\normalfont\ttfamily\color{RawSienna}}%
\makeatother

\begin{document}
\maketitle

\vskip-1.75\baselineskip

{\footnotesize
\halign{\tabskip1em\bfseries#\unskip&\hfil(#\unskip)\hfil&#\unskip\hfil\cr
0.9  & 2022/09/22 & initial release\cr
0.9a & 2022/09/23 & doc fixes, mention \ctanpackage{longfbox}\cr
0.9b & 2022/10/01 & - fix "placement of titled-frame title
                                  is not backwards compatible" (\ghissue{2})\cr
\omit&\omit& - |\colorframedTitleBarFrame| as workaround to upstream issues with titled-frame
               (\ghissue{3}, \ghissue{4})\cr
\omit&\omit& - add link to \ghissue{5} for usage with list items, other doc improvements
               such as \ghissue{1}\cr
}}

\vskip-\baselineskip\vskip0pt\relax

\section{Description}


This package fixes various colour leaks one encounters with
the environments from Donald Arseneau's package
\ctanpackage{framed}.  Typically, colour leaks occur if using
|\color| (at top level or involving a page break) inside the
environments, or more subtly also when using |\textcolor| with
an argument ending up being split at a page break.

This latter type of colour leak (or colour disappearance) is the
more challenging one as the fix requires modifications or
replacements not only of some of the \ctanpackage{framed}.sty
macros (such as its |\CustomFBox|, which \colorframed overwrites)
but also to some \LaTeX{} internals, as some environments of
\ctanpackage{framed}.sty rely on usage of |\fbox| or |\colorbox|.
Rather than overwriting internal \LaTeX{} macros such as
|\@frameb@x| or |\color@b@x|, \colorframed simply replaces |\fbox|
and |\colorbox| in the \ctanpackage{framed}.sty environments by
appropriate substitutes.

I am aware \ctanpackage{tcolorbox} package documentation explains at
least one colour issue which looks similar to those fixed
here in \ctanpackage{framed} context, and that the fix overthere uses
an extra colour stack, hence is not \XeLaTeX{} compatible
currently.

The problems are solved here without involving an extra
colour stack, hence the fixes work also with \XeLaTeX.

\section{The environments from \ctanpackage{framed}}

  We refer the reader to \ctanpackage{framed} for the official
  documentation and provide here only a few relevant details,
  particularly
  regarding the `titled-frame' environment which is described
  in \ctanpackage{framed} documentation more as being a
  template than a user-level finalized environment.
\begin{titled-frame}{A list of the environments from package \ctanpackage{framed}}
  This list indicates which boxing macros are
  used in the original, and their replacement.
% (investigate later why something such as shortverb |#| in
% preamble does not work)
\halign{\bfseries#\hfil&(#\unskip)~\hfil&#\hfil\cr
%
  framed & |\fbox| & Replaced by specialization of own \rlap{\expandafter|\string\CustomFBox|.}\cr
%
  oframed & |\CustomFBox| & \colorframed uses a redefined |\CustomFBox|.\cr
%
  titled-frame & |\CustomFBox| & \emph{id.}\cr
%
  shaded & |\colorbox| & Replaced by |\colorframedcolorbox|.\cr
%
  shaded* & |\colorbox| & \emph{id.}\cr
%
  snugshade & |\colorbox| & \emph{id.}\cr
%
  snugshade* & |\colorbox| & \emph{id.}\cr
%
  leftbar & none & This one does not use any boxing macro.\cr
}
\end{titled-frame}
The previous box gives an example of use of `titled-frame' environment.
  It is an environment with
  one mandatory argument which provides the title of the
  frame, which is repeated after a page break with |(cont)|
  appended.  The colours |TFFrameColor| and
  |TFTitleColor| must have been defined by the user.
  \colorframed adds the possibility to colorize separately the continuation
  label (else it inherits the colour of the frame).  For example we used
\begin{verbatim}
\renewcommand\colorframedTFconlabcolorcommand{\color{purple}}
\end{verbatim}
  and the default definition of this macro is to do nothing.

\medskip
% check why topsep influences separation of item 1 and 2
% it shouldn't? hmm... ah yes I think framed itself uses \topsep...
%\begin{enumerate}[noitemsep,topsep=6pt,font=\normalcolor]
% j'utilise enumitem+noitemsep sinon je vois dans la partie ombrée
% de l'espace supplémentaire au début de l'item (peut-être aussi
% lié à l'emploi du package parskip)
% \colorlet{shadecolor}{lightgray!25}

\begin{footnotesize}
\begin{itemize}[noitemsep,topsep=3pt,% influences strangely item separation
                         parsep=3pt]
\begin{shaded}
\item
\emph{The current box is an example of \emph{`shaded'}
    environment; The
    |shadecolor| must have been defined by user. See the
    \ghissue[repository issue ]{5} for some comments about how to use
    such environment with list items, if the environment is to start
    at start of the item. }
\end{shaded}

\begin{snugshade}
\item \emph{This is an example of \emph{`snugshade'}
    environment. It shares with \emph{`shaded'} the usage of |shadecolor|.}
\end{snugshade}

\begin{snugshade*}
\item \emph{This is an example of \emph{`snugshade*'}.}
\end{snugshade*}

\begin{snugshade}
\item
%
  The `titled-frame' environment was in effect broken in recent \LaTeX\ which
  has modified how |\smash| behaves (the continuation label created a blank
  line interrupting the framing); \colorframed fixes this infelicity in
  passing.

  To customize
  further usage of `titled-frame' one must renew its definition,
  which is left untouched by \colorframed.
  Here is its source from \ctanpackage{framed}.sty: (code and
  comments by Donald Arseneau)%
\begin{verbatim}
% A particular type of titled frame with continuation marks.  
% Parameter #1 is the title, repeated on each page.
\newenvironment{titled-frame}[1]{%
  \def\FrameCommand{\fboxsep8pt\fboxrule2pt
     \TitleBarFrame{\textbf{#1}}}%
  \def\FirstFrameCommand{\fboxsep8pt\fboxrule2pt
     \TitleBarFrame[$\blacktriangleright$]{\textbf{#1}}}%
  \def\MidFrameCommand{\fboxsep8pt\fboxrule2pt
     \TitleBarFrame[$\blacktriangleright$]{\textbf{#1\ (cont)}}}%
  \def\LastFrameCommand{\fboxsep8pt\fboxrule2pt
     \TitleBarFrame{\textbf{#1\ (cont)}}}%
  \MakeFramed{\advance\hsize-20pt \FrameRestore}}%
%  note: 8 + 2 + 8 + 2 = 20.  Don't use \width because the frame title
%  could interfere with the width measurement.
 {\endMakeFramed}
\end{verbatim}
  The continuation label used in the `titled-frame' environment
  is |$\blacktriangleright$|
  (which despite its name reacts to current text colour setting).
  It requires to
  the best of the author knowledge loading \ctanpackage{amssymb} or
  some other math symbols package and it is up to user to do it.
\end{snugshade}

% \begin{shaded} % nesting does not work
\begin{framed}
\item
\emph{(one can not nest environments of \ctanpackage{framed}.sty, else we would
  have done so here.)}

\textcolor{blue}{(Here is an attempt to use the {`framed'} environment
  around a list item; we can see an unexpected
  induced indentation of the
  paragraph. And, by the way, there is no {`snugframe'} environment which would
  let the border fit not the whole width but only the indented item.)}
\end{framed}
% \end{shaded} % nesting does not work

\begin{snugshade}
\item To customize the `titled-frame' output, one may simply redefine the
  environment via suitable changes e.g. to the setting of |\fboxsep| and
  |\fboxrule|.  But doing so, one quickly discovers that the layout from
  |\TitleBarFrame| has problems (see on this the repository issues \ghissue{3}
  and \ghissue{4}).  So \colorframed |0.9b| provides
  |\colorframedTitleBarFrame|, which can be used as a replacement to
  |\TitleBarFrame| in a redefined custom `titled-framed' environment.  Check
  the source code for comments.  In particular this variant
  obeys added customizing
  macros whose defaults are shown here:
\begin{verbatim}
% (such \fboxsep should be set by environment itself, like titled-frame does)
\def\colorframedTFtitlesep{\fboxsep}% hor. distance of title from left border
\def\colorframedTFconlabsep{3pt}% hor. distance of continuation label from border
\end{verbatim}
  This variant handles differently than original the horizontal and vertical
  placement of the title.  It is provided as an exception
  to the general rule that this package should take care only of fixing colour
  problems (addition of \emph{new} environments may be considered in future,
  but fixing existing environments will not be done beyond what is described here).

\end{snugshade}
\end{itemize}
  Despite the fact already
  mentioned that anyhow `titled-frame' had been broken some years ago by an
  upstream \LaTeX\ change we did not change it beyond
  fixing this, and the alternative |\colorframedTitleBarFrame| must be
  opted-for explicitly by user in a re-definition of `titled-frame'
  environment, the default remaining with the problematic
  |\TitleBarFrame|. Or the user can do |\let\TitleBarFrame\colorframedTitleBarFrame|.
  \par
\end{footnotesize}



% The aim of \colorframed regarding \ctanpackage{framed} existing
% code base is strictly limited to fixing the colour leak issues,
% there is no general intent to modify the existing environments of
% \ctanpackage{framed} for them to acquire additional capabilities and
% customizability.

Already we have mentioned
\begin{verbatim}
\colorframedTFconlabcolorcommand
\end{verbatim}
whose default definition let it expands to nothing at all (the continuation label
then inherits the |TFFrameColor| colour).

Regarding the `framed' environment, \colorframed adds the
\begin{verbatim}
\colorframedbordercolorcommand
\end{verbatim}
which expands by default to |\normalcolor|.  It influences the colour of the
framing done by environment `framed', the original having no such
customizability.  Do not redefine it to do nothing, it always should set some
colour, else the borders may display colour loss after a page break.

For example
\begin{verbatim}
\setlength{\FrameRule}{5pt}
\setlength{\FrameSep}{9pt}
\renewcommand\colorframedbordercolorcommand{\color{gray!50}}
\end{verbatim}
configures the next usage of `framed'.

\setlength{\FrameRule}{5pt}
\setlength{\FrameSep}{9pt}
\renewcommand\colorframedbordercolorcommand{\color{gray!50}}
\begin{framed}
  This environment allows indeed customization of the border width
  and of the separation with contents, via |\FrameRule| and
  |\FrameSep|.

  The length |\FrameSep| influences also `oframed', `shaded', and `shaded*',
  but `snugshade' and `snugshade*' employ |\fboxsep| rather.

  |\FrameRule| influences both `framed' and `oframed'.
\end{framed}

\section{Technicalities, musings about future}

\begin{snugshade}
The analysis and original workarounds for using \ctanpackage{framed}.sty with
colours were developed by me in some contributions I made to
the \href{https://github.com/sphinx-doc/sphinx}{Sphinx project} and I am
transferring here the general idea (in much simplified form as the context
overthere is with rendering code snippets using
\ctanpackage{fancyvrb} and
the \href{https://githubcom/pygments/pygments}{Pygmentize} library, and the
solved problems include in particular the automatic wrapping of long code
lines, in a way allowing page-breaks even in the middle of highlighted
strings).

The key thing is that the boxes handled by \ctanpackage{framed}.sty may
contain isolated colour push or colour pop.  We must make
sure an isolated colour push, if followed by colour changes,
is always followed by paired ones, and never by a colour pop
from a colour command originated "prior".

This means we can not ever use things such as |\fbox|, |\colorbox| or
|\fcolorbox| (whether or not \ctanpackage{xcolor} is loaded) as they will
cause colour pops after a possibly isolated colour push contained in the
(already boxed) contents which are fetched to them as arguments by
\ctanpackage{framed}.sty internals.  We must re-implement them otherwise.
This is what is done here, without touching the \LaTeX\ internal macros
themselves.

Apart from that, part of the macro file is occupied with the
`titled-frame' environment of \ctanpackage{framed}.sty, which was broken by an
upstream LaTeX change related to |\smash|, and has obvious problems
apart from that.

Further "obvious problems" do exist for the other \ctanpackage{framed}.sty
environments but it is not the matter of this package to fix them.
Perhaps in future another package will be provided based upon the
work here and adding a completely new environment fixing the "obvious
problems" (some of them being simply lack of customizability).

The author has developed based upon usage of \ctanpackage{pict2e}
breakable boxes with round corners, background colour, optional
shadow (possibly inset), and other goodies and was planning on
incorporating this environment into the package at some time in future.

% Hopefully this addition will be done when time will permit.

% \thispagestyle{empty}
% \enlargethispage{2\baselineskip}

%\colorlet{shadecolor}{lightgray!50}
% \footnotesize
  But after initial release made it to CTAN on
  2022/09/22 I became aware of \ctanpackage{longfbox} which
  provides already such \ctanpackage{pict2e} breakable boxes with
  rounded corners (even elliptical arcs), and furthermore with a
  CSS-like interface which is exactly what I had done on my side
  too... I need to check more \ctanpackage{longfbox} before a
  decision is made here!  Perhaps it will be better to keep
  \colorframed as it is currently and
  extras such as new \ctanpackage{pict2e}-based boxes with a key-value
  interface to `inline' or `display' boxing macros and environments
  should make it to another package (loading \colorframed of course).
%
Of course, such a package would remain limited in comparison to the fully
customizable boxes provided by package \ctanpackage{tcolorbox} (ot
those of \ctanpackage{mdframed}) but our testing showed
significant speed-up in build time, which may matter for long
documents.

\end{snugshade}
\centerline{\hrulefill documentation ends here\hrulefill}
\end{document}